---
title: "Titanic"
author: "Tal Barami"
date: "24 בנובמבר 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Set working directory:
```{r}
knitr::opts_knit$set(root.dir = 'C:\\Users\\Tal\\Desktop\\Titanic\\Titanic')
getwd()
```

Read the train.csv file into a dataframe. 
```{r}
df <- read.csv("Titanic/train.csv",na.strings = "")
```


Check the datatypes of the attributes using the *str* method. 
```{r}
str(df)
```

Convert numeric features into factors.
```{r}
df$Survived<- as.factor(df$Survived)
df$Pclass<- as.factor(df$Pclass)
```

Produce some descriptive statistics using the *Summary* method, and read it to get to know your data. Do you have missing data(NA) inside? If so, you should either preprocess the data to have non NAs, or use only models that can handle missing values. We will go with the second option, so you don't need to preprocess the data for now.
```{r}
summary(df)
```

It seems that the following features have too many categories or simply represent an index: PassengerId, Ticket, Name. We'll ignore these features.
```{r}
df_no_cabin <- df[,-c(1,4,9, 11)]
df <- df[,-c(1,4,9)]
```


Divide the features' names to numerics and factors:
```{r}
cols<- 1:dim(df)[2]
factors <- cols[sapply(df,is.factor)]
numerics <- cols[!sapply(df,is.factor)]
```

#Many algorithms deal with numerical input only.
Convert our data to a numeric form, so we can examine random forest:
```{r}
df.numeric <- df_no_cabin
df.numeric$Survived=as.numeric(df_no_cabin[,'Survived'] == "1")
```

Use CARET's *dummyVars* to convert nominal features to various dummies or flags, according to each feature's possible levels.
```{r}
library(caret)
dmy <- dummyVars(" ~ .", data = df.numeric,fullRank=F)
df.numeric <- data.frame(predict(dmy, newdata = df.numeric))
head(df.numeric)
```

```{r}
df.numeric$Survived<- as.factor(df.numeric$Survived)
levels(df.numeric$Survived)<-c("x0","x1")
```

Omit records that contain empty values:
```{r}
empty_records <- df.numeric[rowSums(is.na(df.numeric)) > 0,]
if (nrow(empty_records)>0)
  df.numeric<- na.omit(df.numeric)
```

Tide the data two times: the first is for categorial data and the second for numeric data.
```{r}
#install.packages("tidyr")
library(tidyr)
df_tidy_factors<-gather(df[,factors],"feature","value",-1)
df_tidy_numerics<-gather(cbind(Survived=df[,1],df[,numerics]),"feature","value",-1)
```


The first plot describes only categorical features (factors). 
Notice that the *scales* parameter was set to "free" to enable a suitable scaling for each facet (otherwise it is hard to view some of the facets, that need much smaller scales). We use the *facet_grid* that accepts a *scales* parameter.

```{r}
#install.packages("ggplot2")
library(ggplot2)
qplot(x=value,data=df_tidy_factors,fill=Survived) + facet_grid(~feature,scales="free")
```

One more plot for numeric features:
```{r}
qplot(x=value,data=df_tidy_numerics,fill=Survived) + facet_grid(~feature,scales="free")
```

It certainly looks like there are more chances to survive in certain levels of almost each feature.

Split the data into 80% train and 20% test sets.
```{r}
indices <- sample(1:nrow(df),nrow(df)*0.7)
train<- df[indices,]
test<- df[-indices,]
```

# Train a naive bayes model:
# Train a Random Forest model, evaluating it with ROC.
```{r}
set.seed(123)
fit.rf <- train(Survived~., data=df.numeric, method="rf",ntree = 10,  
              metric='ROC', 
              trControl=trainControl(method="cv", 
                                         number=5,
                                         savePredictions="final",
                                         summaryFunction = twoClassSummary,
                                         classProbs = TRUE),
              tuneGrid=expand.grid(.mtry=c(5)),
              na.action = na.pass)
```


# Train an xgb tree model:
```{r}
library(caret)
library(lattice)
library(xgboost)


set.seed(123)
control <- trainControl(method="cv", number=5)

grid <- expand.grid(.nrounds=20,.max_depth=6,.eta=c(0.1,0.3,0.5),.gamma=0.1,.colsample_bytree=0.5,.min_child_weight=0.01,.subsample=0.7)

fit.xgb <- train(Survived~., data=df, method="xgbTree", metric="Accuracy", trControl=control, verbose=FALSE,tuneGrid =grid, na.action=na.pass)

plot(fit.xgb)
```
Predict the target feature for records of the test set, using the trained model.
```{r}
xgb_pred <- predict(fit.xgb, test, na.action = na.pass)
```

Produce a confusion matrix between predicted target attributes and real target attributes of the test set.
```{r}
table(xgb_pred,test$Survived)
```

Calculate the model's test accuracy.
```{r}
mean(xgb_pred==test$Survived)
```


#Train a gbm decision-tree model based on the train data.
```{r}
library(caret)
library(lattice)
library(gbm)

control <- trainControl(method="cv", number=5)
grid <- expand.grid(.n.trees=c(5),.interaction.depth=6,.shrinkage=c(0.1,0.3,0.5),.n.minobsinnode=5)
set.seed(123)
fit.gbm <- train(Survived~., data=df, method="gbm", metric="Accuray", trControl=control, verbose=FALSE,tuneGrid =grid, na.action=na.pass)
plot(fit.gbm)
```

Predict the target feature for records of the test set, using the trained model.
```{r}
gbm_pred <- predict(fit.gbm, test, na.action = na.pass)
```

Produce a confusion matrix between predicted target attributes and real target attributes of the test set.
```{r}
table(gbm_pred,test$Survived)
```

Calculate the model's test accuracy.
```{r}
mean(gbm_pred==test$Survived)
```


### Results:
Load the test file.
```{r}
new_df <-read.csv('Titanic/test.csv',na.strings = "")
```

Create a vector with the PassengerIds of records in the test file. We will soon attach it to the prediction results.
```{r}
ids<- new_df$PassengerId
```


Repeat the same preprocessing steps that were performed on the train data (factorizing two features and ignoring three features). Train and test data must have the same structure (except for the target feature)
```{r}
new_df$Pclass<- as.factor(new_df$Pclass)
new_df_no_cabin<- new_df[,-c(1,3,8,10)]
new_df<- new_df[,-c(1,3,8)]

new_df.numeric <- new_df_no_cabin
new_dmy <- dummyVars(" ~ .", data = new_df.numeric,fullRank=F)
new_df.numeric <- data.frame(predict(new_dmy, newdata = new_df.numeric))
head(new_df.numeric)
empty_records <- new_df.numeric[rowSums(is.na(new_df.numeric)) > 0,]
if (nrow(empty_records)>0)
  new_df.numeric<- na.omit(new_df.numeric)
```

Another thing that is going to make troubles is that the test data contains some new levels in the *Cabin* feature that did not appear in the train data.

- Getting the levels of the "Cabin" feature from a model *m*: m$xlevels[["Cabin"]]

- Getting the levels of the "Cabin" feature from a dataframe *d*: levels(d$Cabin)

- Union operation: union(x,y)


```{r}
fit.xgb$xlevels[["Cabin"]] <- union(fit.xgb$xlevels[["Cabin"]], levels(new_df$Cabin))
fit.gbm$xlevels[["Cabin"]] <- union(fit.gbm$xlevels[["Cabin"]], levels(new_df$Cabin))
```


Load the test file and predict the target attribute for its records using your trained model. Don't forget to set *na.action* to "na.pass" if you used caret in the model creation, otherwise it will fail.

```{r}
new_xgb_pred<- predict(fit.xgb,new_df,na.action = na.pass)
new_gbm_pred<- predict(fit.gbm,new_df,na.action = na.pass)
new_rf_pred<- predict(fit.rf,new_df.numeric,na.action = na.pass)
```


19. Write the *PassengerId* and *Survived* attributes to a csv file and submit this file to kaggle's competition (You should first subscribe to the site). What is your score?

```{r}
xgb_res <- cbind(PassengerId=ids,Survived=as.character(new_xgb_pred))
write.csv(xgb_res,file="Titanic/xgb.csv",row.names = F)

gbm_res <- cbind(PassengerId=ids,Survived=as.character(new_gbm_pred))
write.csv(gbm_res,file="Titanic/gbm.csv",row.names = F)

rf_res <- cbind(PassengerId=ids,Survived=as.character(new_rf_pred))
write.csv(rf_res,file="Titanic/rf.csv",row.names = F)
```

